Merging data
========================================================

It's often useful to be able to merge multiple datasets that share a common identifier.  We've used this to ask questions which span multiple topics in the GapMinder.org collection of datasets.  

In this example, we pull together a dataset with traffic mortality rate in 2004 as well as deaths from falls in 2004.  This utilizes the material described in chapter 12 of the "Compendium" (Data Management).

```{r message=FALSE}
require(mosaic)
trellis.par.set(theme=col.mosaic())
options(digits=3)
```

Read the traffic dataset (and clean it up)

> traffic deaths per 100,000

```{r}
ds1 = read.csv("http://www.math.smith.edu/~nhorton/traffic.csv")
names(ds1)
names(ds1)[names(ds1)=="Traffic.mortality.per.100.000..age.adjusted"] = "country"
names(ds1)[names(ds1)=="X2004"] = "traffic2004"
ds1 = subset(ds1, select=c("country", "traffic2004"))
head(ds1)
```

Read the falls dataset (and clean it up)

> mortality from falling per 100,000
```{r}
ds2 = read.csv("http://www.math.smith.edu/~nhorton/falls.csv")
names(ds2)
names(ds2)[names(ds2)=="Falls..age.adjusted.mortality.per.100.000"] = "country"
names(ds2)[names(ds2)=="X2004"] = "falls2004"
ds2 = subset(ds2, select=c("country", "falls2004"))
head(ds2)
```

Merge datasets and take a look at the results.
```{r}
ds = merge(ds1, ds2, by="country")
head(ds)
```

Let's peruse the datasets and look for interesting observations.

```{r}
favstats(~ falls2004, data=ds)
favstats(~ traffic2004, data=ds)
subset(ds, traffic2004 > 115)
subset(ds, falls2004 > 17)
subset(ds, is.na(traffic2004))
noiraq = subset(ds, country != "Iraq")
```

Unadorned scatterplot
```{r fig.width=7, fig.height=6}
xyplot(traffic2004 ~ falls2004, data=ds)
```

Scatterplot displaying the country names. (Note use of "with()" to allow the "panel.labels()" function to access the country variable).

```{r fig.width=7, fig.height=6}
panel.labels = function(x, y, col='black', labels='x', ...) 
   { panel.text(x, y, labels, col=col, ...)}
with(noiraq, xyplot(traffic2004 ~ falls2004, cex=0.5, 
       panel=panel.labels,
       labels=country))
```


