TeachStatsWithR
===============

```{r include=FALSE}
require(mosaic)
```

Materials for the MOSAIC "Teaching Statistics with R and RStudio"

### Outline

This project consists of several short books that are inter-related.

1. Start Teaching with R
    Directory: `Starting`
2.  The Core of a Traditional Course
    Directory: `Traditional`
3. Simulation-Based Inference
    Directory: `Simulation`
4. Functions and Formulas
    Directory: `Functions`
5. Teaching with Internet Services
    Directory: `Internet`
6. Start with Modeling
    Directory: `Modeling`

Outlines for the individual chapters are in the `Outline.Rmd` file in each directory.

## Overview

Some general comments about the project as a whole.

```{r starting-outline,child='Starting/Outline-Starting.Rmd',eval=TRUE}
```

```{r traditional-outline,child='Traditional/Outline-Traditional.Rmd',eval=TRUE}
```

```{r simulation-outline,child='Simulation/Outline-Simulation.Rmd',eval=TRUE}
```

```{r functions-outline,child='Functions/Outline-Functions.Rmd',eval=TRUE}
```

```{r internet-outline,child='Internet/Outline-Internet.Rmd',eval=TRUE}
```

```{r modeling-outline,child='Modeling/Outline-Modeling.Rmd',eval=TRUE}
```

